Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
revamped intro slides as per Jasons suggestions.

\section{Motivation}

\frame{\frametitle{Single-headedness}
  Dependency parse treebanks today are single-headed.

@@

%\subsection{Multi-headed}
\frame{\frametitle{Multi-headedness}
  Multi-headedness Can Capture Additional Linguistic Phenomenon
  \begin{itemize}
    \item Control
    \item Relativization
    \item Conjunction
  \end{itemize}
}

@@

\frame{\frametitle{Link Grammars: Example Parse}
  \begin{figure}
    \begin{dependency}[edge style={-}]
	  \begin{deptext}
	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  \end{deptext}
	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	  \deproot[edge above, thick, edge style = {red}]{2}{W}
	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
	  \deproot[edge above, thick, edge style = {red}]{10}{X}
	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
    \end{dependency}
    \caption*{Link Parse of a sentence from Penn Tree Bank}
  \end{figure}
}

\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
  \begin{figure}
    \begin{dependency}
	  \begin{deptext}
	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  \end{deptext}
	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	  \deproot[edge above, thick, edge style = {red}]{2}{W}
	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
	  \deproot[edge above, thick, edge style = {red}]{10}{X}
	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
    \end{dependency}

@@

>>>>

\section{Introduction}
\frame{\frametitle{Introduction}
\begin{itemize}
\item This talk is about automatically converting from one annotation style to another.
\item The conversion could be hard, where information is fragmented or missing.

%This is often useful.  You may want to do it yourself.  
\item We use a general technique, Integer Linear Programming to recover this missing information.
\end{itemize}
}

\frame{\frametitle{In Our Case: What We Started With}
  \begin{figure}
    \begin{dependency}[edge style={-}]
	  \begin{deptext}
	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  \end{deptext}
	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	  \deproot[edge above, thick, edge style = {red}]{2}{W}
	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
	  \deproot[edge above, thick, edge style = {red}]{10}{X}
	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
    \end{dependency}
    \caption*{Link Grammar: Parse with undirected edges}
  \end{figure}
}

\frame{\frametitle{What We Want:}
  \begin{figure}
    \begin{dependency}
	  \begin{deptext}
	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  \end{deptext}
	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	  \deproot[edge above, thick, edge style = {red}]{2}{W}
	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
	  \deproot[edge above, thick, edge style = {red}]{10}{X}
	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
    \end{dependency}
    \caption*{Multiheaded parse with directionalized edges}
  \end{figure}

\frame{\frametitle{Why We Want That}
  \begin{itemize}
    \item We want to develop parsing algorithms for parses that look like this
    \item We couldn't figure out where to get the data to test them.
  \end{itemize}
}

\frame{\frametitle{Single-headedness}
  \begin{itemize}
  \item Dependency parse treebanks today are either single-headed or not planar.
  \item Stanford Dependencies are multiheaded but not planar
  \end{itemize}

@@

\uncover<2->{\begin{center}
    Link Grammar is almost a multiheaded planar corpora! We just need to directionalize the links.
\end{center}}
}

\section{Motivation}
%\subsection{Multi-headed}
\frame{\frametitle{Why Multi-headedness?}
  Multi-headedness Can Capture Additional Linguistic Phenomenon
  \begin{itemize}
    \item Control
    \item Relativization
    \item Conjunction
  \end{itemize}
}

@@

\frame{\frametitle{Link Grammars: Same Example Parse Again}
  \begin{figure}
    \begin{dependency}[edge style={-}]
	  \begin{deptext}
	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  \end{deptext}
	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	  \deproot[edge above, thick, edge style = {red}]{2}{W}
	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
	  \deproot[edge above, thick, edge style = {red}]{10}{X}
	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
    \end{dependency}
    \caption*{Link Parse of a sentence from Penn Tree Bank}
  \end{figure}
}

\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph Again}
  \begin{figure}
    \begin{dependency}
	  \begin{deptext}
	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  \end{deptext}
	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	  \deproot[edge above, thick, edge style = {red}]{2}{W}
	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
	  \deproot[edge above, thick, edge style = {red}]{10}{X}
	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
    \end{dependency}

@@

\frame{\frametitle{Link Results: Subject-Verb links are backwards}
\begin{itemize}
      %\item Different dependency formalisms disagree on whether a clause is headed by the aux verb or the main verb.
  \item The Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. 
  \item Sometimes the governing verb links to the auxilliary, and sometimes to the main, depending on the type of clause. 
  \item But the governing verb usually links to the subject when there is one.
  \item So this makes the subject a consistent choice to make the head of a clause.
\end{itemize}

%governing verb -\> subject -\> auxilliary verb -\> main

\begin{center}
  To fix this, we could edit the link grammar, link parses, or the ILP. %\\ Our technique is only as useful as the input given.
\end{center}

    %Question: Which do we pick?
    %Answer: Neither! We actually pick the subject as the head of the clause! 
    %Question: Why??
    %Answer:
    %    The LG folks seem to be inconsistent within their grammar about whether the aux verb or the main verb is the head of a clause. 
    %        Sometimes the governing verb links to aux, and sometimes to main, depending on the type of clause.
    %    But the governing verb usually links to the subject when there is one!
    %    Therefore, the subject ends up being a consistent choice for the head of the clause.  We get governing_verb --> subject --> aux --> main as our annotation scheme.
    %Question: Couldn't this be fixed?
    %Answer: Of course it could.  You could edit the link grammar or the link parses or the ILP.  Our technique is only as useful as the input you give it ...  
}
